---
title: "Expoloratory analysis for Pilot-B data"
author: "Anjie Cao (anjiecao@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    code_folding: hide
    toc_depth: 3
    toc_float:
      collapsed: true
---



```{r include=FALSE}
library(tidyverse)
library(here)
library(rstatix)
library(lme4)
library(DT)
library(jsonlite)
library(tidyverse)
library(kableExtra)
source(here("script/00_read_raw_data.R"))
source(here("script/01_apply_exclusion_criteria.R"))
source(here("script/02_tidy_data.R"))

RAW_DATA_DIR <- here("raw_data/")
AGGREGATED_DATA_PATH <- here("processed_data/aggregated_data.csv")
PROCESSED_DATA_PATH <- here("processed_data/processed_data.csv")
NO_FILTER_DATA_PATH <- here("no_filtered.csv")
####Aggregate across participants, clean up demographic fields
aggregate_data(RAW_DATA_DIR, AGGREGATED_DATA_PATH)


####Import data 
data_with_demog <- read_csv(AGGREGATED_DATA_PATH)

#### Data exclusion / filtering
# participants level exclusion
filtered_data <- filter_demog(data_with_demog)
filtered_data <- filter_spacebar(filtered_data)
filtered_data <- filter_trial_variation(filtered_data)
filtered_data <- filter_too_many_outliers(filtered_data)
#trial level exclusion
filtered_data <- filter_outlier_trial(filtered_data)  

#### Prepare data for analysis - create columns etc.
d <- tidy_data(filtered_data,PROCESSED_DATA_PATH) # this also writes

```

# Visualizing raw trial looking time
```{r}
d_no_target <- d %>% 
  filter(trial_stimulus_type != "target")
```


## Looking time distribution {.tabset}
### RAW 
```{r fig.height=10, fig.width=10}
data_with_demog %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) + 
  xlim(0, 9000)

data_with_demog %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) + 
  xlim(0, 9000) + 
  facet_wrap(~subject)
```


### AFTER EXCLUSION 
```{r fig.height=10, fig.width=10}
d %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) + 
  xlim(0, 6000)

d %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) + 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### Exclude the target trial
```{r}
d_no_target %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) + 
  xlim(0, 6000)

d_no_target %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) + 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### complexity difference
```{r}
d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### block difference
```{r}
d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### Complexity differences no target 
```{r}
d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### Block differences no target
```{r}
d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)

```

# Visualizing aggregarded looking time {.tabset}
```{r}
d_sum_individual <- d %>% 
  group_by(subject, block) %>% 
  summarise(
    mean_lt = mean(trial_looking_time, na.rm = TRUE), 
    sd = sd(trial_looking_time, na.rm = TRUE), 
    n = n(), 
    ci_range_95 = qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)), 
    ci_ub = mean_lt + ci_range_95, 
    ci_lb = mean_lt - ci_range_95
  )

d_sum <- d %>% 
  group_by(block) %>% 
  summarise(
    mean_lt = mean(trial_looking_time, na.rm = TRUE), 
    sd = sd(trial_looking_time, na.rm = TRUE), 
    n = n(), 
    ci_range_95 = qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)), 
    ci_ub = mean_lt + ci_range_95, 
    ci_lb = mean_lt - ci_range_95
  )
```

## aggregated 
```{r}
d_sum %>% ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

## individual 
something weird happened
```{r}
d_sum_individual %>% 
  ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
  

```


```{r fig.width=15, fig.height=15}
d_sum_individual %>% 
  ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub))  + 
  facet_wrap(~subject) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
  
```

## weird ones
We can see three very weird ones: 
SS1604513317537
SS1604515995769
SS1604516882157
 
```{r}
weird <- c("SS1604513317537", 
           "SS1604515995769", 
           "SS1604516660396")

d_sum_individual %>% 
  filter(subject %in% weird) %>% 
  ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub))  + 
  facet_wrap(~subject) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

d %>% 
  filter(subject %in% weird) %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) +
  facet_wrap(~subject) + 
  xlim(0, 6000)+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

## Excluding weird ones 
```{r}
d_no_weird <- d %>% 
  filter(!(subject %in% weird)) 

  
d_no_weird_sum <- d_no_weird %>% 
  group_by(block) %>% 
  summarise(
    mean_lt = mean(trial_looking_time, na.rm = TRUE), 
    sd = sd(trial_looking_time, na.rm = TRUE), 
    n = n(), 
    ci_range_95 = qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)), 
    ci_ub = mean_lt + ci_range_95, 
    ci_lb = mean_lt - ci_range_95
  )

d_no_weird_sum %>% ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
# Basic model 
```{r}
null_m <- lmer(trial_looking_time ~ 1 + (1|subject), 
               data = d)

basic_m <- lmer(trial_looking_time ~ block + (1|subject), 
     data = d)

anova(null_m, basic_m)
```



# TODO: 
- age, SES, gender

# number of background repetition on looking time {.tabset}
## full, aggregated, by complexity
```{r}
full_aggregated <- d %>% 
  mutate(
    number = 1) %>% 
  group_by(
  subject, block, trial_stimulus_type
  ) %>% 
  mutate(num_times_stimulus_seen = cumsum(number))

full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = trial_stimulus_complexity)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```

## full, aggregated, by blocks 
```{r}
full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = block)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```


## no weird, aggregated, by complexity
```{r}
excluded_sum <- full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  filter(!(subject %in% weird))

excluded_sum %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = trial_stimulus_complexity)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```

## no weird, aggregated, by blocks
```{r}
excluded_sum %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = block)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```

## individual, by complexity 
```{r}
full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = trial_stimulus_complexity)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10)) + 
  facet_wrap(~subject)
```

## individual, by block 
```{r}
full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = block)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10)) + 
  facet_wrap(~subject)
```

## models 
```{r}
null_m <- lmer(log(trial_looking_time) ~ 1 + (1|subject), 
               data = excluded_sum)

rep_m <- lmer(log(trial_looking_time) ~ num_times_stimulus_seen + (1|subject), 
     data = excluded_sum)

interaction_m <- lmer(log(trial_looking_time) ~ num_times_stimulus_seen * block + (1|subject), 
     data = excluded_sum)
anova(rep_m, null_m)
anova(rep_m, interaction_m)
summary(rep_m)
```






# target reaction time predictive of anything? 



# import the complexity norm 

