---
title: "Expoloratory analysis for Pilot-B data"
author: "Anjie Cao (anjiecao@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    code_folding: hide
    toc_depth: 3
    toc_float:
      collapsed: true
---

# todo: 
- new exclusion criteria 
- ideal learner stuff 
- look at target reaction time? 
- import the complexity norm? 


```{r include=FALSE}
library(tidyverse)
library(here)
library(rstatix)
library(lme4)
library(DT)
library(jsonlite)
library(tidyverse)
library(kableExtra)
source(here("script/00_read_raw_data.R"))
source(here("script/01_apply_exclusion_criteria.R"))
source(here("script/02_tidy_data.R"))

RAW_DATA_DIR <- here("raw_data/")
AGGREGATED_DATA_PATH <- here("processed_data/aggregated_data.csv")
PROCESSED_DATA_PATH <- here("processed_data/processed_data.csv")
NO_FILTER_DATA_PATH <- here("no_filtered.csv")
####Aggregate across participants, clean up demographic fields
aggregate_data(RAW_DATA_DIR, AGGREGATED_DATA_PATH)


####Import data 
data_with_demog <- read_csv(AGGREGATED_DATA_PATH)

#### Data exclusion / filtering
# participants level exclusion
filtered_data <- filter_demog(data_with_demog)
filtered_data <- filter_spacebar(filtered_data)
filtered_data <- filter_trial_variation(filtered_data)
filtered_data <- filter_too_many_outliers(filtered_data)
#trial level exclusion
# filtered_data <- filter_outlier_trial(filtered_data)  

#### Prepare data for analysis - create columns etc.
d <- tidy_data(filtered_data,PROCESSED_DATA_PATH) # this also writes

```

# Visualizing raw trial looking time
```{r}
d_no_target <- d %>% 
  filter(trial_stimulus_type != "target")
```


## Looking time distribution {.tabset}
### RAW 
```{r }
data_with_demog %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) + 
  xlim(0, 9000)
```
```{r }
data_with_demog %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram() + 
  scale_x_log10()
```


### AFTER EXCLUSION 
```{r }
d %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram() + 
  scale_x_log10()
# 
# d %>% 
#   ggplot(aes(x = trial_looking_time)) + 
#   geom_histogram(bins = 90) + 
#   xlim(0, 6000) + 
#   facet_wrap(~subject)
```

# Digging through the second hump

* Something wacky is happening in background trials.
* Deviant trials are slower
* target trials are likely a mixture between hits and misses

```{r}
d %>% 
  ggplot(aes(x = trial_looking_time, 
             fill = trial_pressed_space_bar)) + 
  geom_histogram() + 
  scale_x_log10() + 
  facet_wrap(~trial_stimulus_type)
```

Zoom in on background trials. The "hump" (second mode) is there in all parts of the experiment.

```{r}
d %>% 
  mutate(block_num = ceiling(total_trial_num / 48)) %>%
  filter(trial_stimulus_type == "background") %>%
  ggplot(aes(x = trial_looking_time, 
             fill = trial_stimulus_complexity)) + 
  geom_histogram() + 
  scale_x_log10() + 
  facet_wrap(~block_num)
```

```{r}
d %>% 
  filter(trial_stimulus_type == "background") %>%
  group_by(subject) %>%
  mutate(mlog = mean(log(trial_looking_time)), 
         sdlog = sd(log(trial_looking_time)),
         outlier = log(trial_looking_time) > mlog + 3*sdlog |
           log(trial_looking_time) < mlog - 3*sdlog) %>%
  ggplot(aes(x = trial_looking_time, fill = outlier)) + 
  geom_histogram() + 
  scale_x_log10() + 
  facet_wrap(~subject)
```

We see this pattern within subs to some extent, irrespective of outliers. Is it after target press?

No, it seems that even on background trials where there was not a target beforehand.

```{r}
d %>% 
  group_by(subject) %>%
  mutate(t_to_b = c(FALSE, diff(as.numeric(as.factor(trial_stimulus_type == "target"))) == -1) & trial_stimulus_type == "background",
         d_to_b = c(FALSE, diff(as.numeric(as.factor(trial_stimulus_type == "deviant"))) == -1) & trial_stimulus_type == "background") %>%
  filter(trial_stimulus_type == "background") %>%
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram() + 
  scale_x_log10() + 
  facet_grid(d_to_b~t_to_b)

```


just look at the best of trials of each individual

```{r}
d %>% 
  group_by(subject) %>% 
  mutate(mlog = mean(log(trial_looking_time)), 
         sdlog = sd(log(trial_looking_time)),
         best_trials = log(trial_looking_time) < mlog + 2*sdlog &&
           log(trial_looking_time) > mlog - 2*sdlog) %>%
  filter(best_trials == TRUE) %>% 
  ggplot(aes(x = trial_looking_time, fill = trial_stimulus_complexity)) + 
  geom_histogram() + 
  scale_x_log10() 
```
## block type by complexity

only looking at trials within 2sd of the means 
```{r}
d %>% 
  group_by(subject) %>% 
  mutate(mlog = mean(log(trial_looking_time)), 
         sdlog = sd(log(trial_looking_time)),
         best_trials = log(trial_looking_time) < mlog + 2*sdlog &&
           log(trial_looking_time) > mlog - 2*sdlog) %>%
  filter(best_trials == TRUE) %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram() + 
  scale_x_log10() +
  facet_wrap(~block)
```

doesn't seem to make a huge difference with or without the outlier 
```{r}
d %>% 
  group_by(subject) %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram() + 
  scale_x_log10() +
  facet_wrap(~block)
```

break down by trial type and stimuli type 

```{r}
d %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram()+
  scale_x_log10() +
  facet_grid(block~trial_stimulus_type)
```

ok, maybe it's time to consider number of repetitions. maybe the second hump means participants getting very bored and they started zoning out???

```{r}
library(mgcv)

rep_d <- d %>% 
  mutate(
    number = 1) %>% 
  group_by(
  subject, block, trial_stimulus_type
  ) %>% 
  mutate(num_times_stimulus_seen = cumsum(number)) %>% 
  filter(trial_stimulus_type == "background")

gam_d <- rep_d %>% 
  mutate(trial_stimulus_complexity = as.factor(trial_stimulus_complexity), 
         block = as.factor(block))

gam_m <- gam(log(trial_looking_time) ~ s(num_times_stimulus_seen), 
             data = gam_d, 
             method = "REML")
summary(gam_m)
plot(gam_m)
```

note this is not a very good model
```{r}
gam.check(gam_m)
```

but still let's give it a try: maybe 10 repetitions are the "turning point" 

```{r}
rep_d %>% 
  mutate(
    reptition_type = if_else(num_times_stimulus_seen > 10, "more", "less")
  ) %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram()+
  scale_x_log10() +
  facet_grid(block~reptition_type)
```
umm this really doesn't make any sense to me...










## age really has nothing to do with it 
```{r}
d %>% 
  mutate(
    m_age = mean(demog_age, na.rm = TRUE), 
    age_group = if_else(demog_age < m_age | demog_age == m_age, "young", "old")
  ) %>% 
  filter(!is.na(age_group)) %>% 
  ggplot(aes(x = trial_looking_time, fill = age_group)) + 
  geom_histogram() + 
  scale_x_log10() 
```



























### complexity difference
```{r}
d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### block difference
```{r}
d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### Complexity differences no target 
```{r}
d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = trial_stimulus_complexity), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)
```

### Block differences no target
```{r}
d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000)

d_no_target %>% 
  ggplot(aes(x = trial_looking_time, 
            fill = block), 
         ) + 
  geom_density(alpha = 0.5)+ 
  xlim(0, 6000) + 
  facet_wrap(~subject)

```

# Visualizing aggregarded looking time {.tabset}
```{r}
d_sum_individual <- d %>% 
  group_by(subject, block) %>% 
  summarise(
    mean_lt = mean(trial_looking_time, na.rm = TRUE), 
    sd = sd(trial_looking_time, na.rm = TRUE), 
    n = n(), 
    ci_range_95 = qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)), 
    ci_ub = mean_lt + ci_range_95, 
    ci_lb = mean_lt - ci_range_95
  )

d_sum <- d %>% 
  group_by(block) %>% 
  summarise(
    mean_lt = mean(trial_looking_time, na.rm = TRUE), 
    sd = sd(trial_looking_time, na.rm = TRUE), 
    n = n(), 
    ci_range_95 = qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)), 
    ci_ub = mean_lt + ci_range_95, 
    ci_lb = mean_lt - ci_range_95
  )
```

## aggregated 
```{r}
d_sum %>% ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

## individual 
something weird happened
```{r}
d_sum_individual %>% 
  ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
  

```


```{r fig.width=15, fig.height=15}
d_sum_individual %>% 
  ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub))  + 
  facet_wrap(~subject) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
  
```

## weird ones
We can see three very weird ones: 
SS1604513317537
SS1604515995769
SS1604516882157
 
```{r}
weird <- c("SS1604513317537", 
           "SS1604515995769", 
           "SS1604516660396")

d_sum_individual %>% 
  filter(subject %in% weird) %>% 
  ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub))  + 
  facet_wrap(~subject) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

d %>% 
  filter(subject %in% weird) %>% 
  ggplot(aes(x = trial_looking_time)) + 
  geom_histogram(bins = 90) +
  facet_wrap(~subject) + 
  xlim(0, 6000)+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

## Excluding weird ones 
```{r}
d_no_weird <- d %>% 
  filter(!(subject %in% weird)) 

  
d_no_weird_sum <- d_no_weird %>% 
  group_by(block) %>% 
  summarise(
    mean_lt = mean(trial_looking_time, na.rm = TRUE), 
    sd = sd(trial_looking_time, na.rm = TRUE), 
    n = n(), 
    ci_range_95 = qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)), 
    ci_ub = mean_lt + ci_range_95, 
    ci_lb = mean_lt - ci_range_95
  )

d_no_weird_sum %>% ggplot(aes(x = block, y = mean_lt)) + 
  geom_pointrange(aes(ymin = ci_lb, ymax = ci_ub)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

# Demographic 
## SES 
education == "Some high school" ~ 1,
        education == "High school diploma" ~ 2, 
        education == "Associate Degree/Technical certification" ~ 3, 
        education == "Bachelor's Degree" ~ 4, 
        education == "Master's Degree" ~ 5, 
        education == "Doctorate/Professional degree" ~ 6, 
        education == "Other" ~ NA_real_
```{r}
d %>% 
  distinct(subject, .keep_all = TRUE) %>% 
  ggplot(aes(x = demog_education)) + 
  geom_histogram()
```
```{r}
null_m <- lmer(trial_looking_time ~ 1 + (1|subject), 
               data = filter(d, !is.na(demog_education)))

edu_m <- lmer(trial_looking_time ~ demog_education + (1|subject), 
     data = filter(d, !is.na(demog_education)))

summary(edu_m)
anova(null_m, edu_m)
```


## Age 

```{r}
d %>% 
  distinct(subject, .keep_all = TRUE) %>% 
  ggplot(aes(x = demog_age)) + 
  geom_histogram()
```
```{r}
null_m <- lmer(trial_looking_time ~ 1 + (1|subject), 
               data = filter(d, !is.na(demog_age)))

age_m <- lmer(trial_looking_time ~ demog_age + (1|subject), 
     data = filter(d, !is.na(demog_age)))

summary(age_m)
anova(null_m, age_m)
```




# number of background repetition on looking time {.tabset}
## full, aggregated, by complexity
```{r}
full_aggregated <- d %>% 
  mutate(
    number = 1) %>% 
  group_by(
  subject, block, trial_stimulus_type
  ) %>% 
  mutate(num_times_stimulus_seen = cumsum(number))

full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = trial_stimulus_complexity)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```

## full, aggregated, by blocks 
```{r}
full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = block)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```


## no weird, aggregated, by complexity
```{r}
excluded_sum <- full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  filter(!(subject %in% weird))

excluded_sum %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = trial_stimulus_complexity)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```

## no weird, aggregated, by blocks
```{r}
excluded_sum %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = block)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10))
```

## individual, by complexity 
```{r}
full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = trial_stimulus_complexity)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10)) + 
  facet_wrap(~subject)
```

## individual, by block 
```{r}
full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  ggplot(
    aes(y = log(trial_looking_time), 
        x = num_times_stimulus_seen,
        color = block)
  ) + 
  geom_point(aes(alpha = 0.2), size = 3, shape = ".") + 
  guides(alpha = FALSE) + 
  labs(color = "Stimulus Complexity") + 
  ylab("Mean Looking Time (log)") + 
  xlab("Number of Stimulus Reptitions") + 
  geom_smooth(method = "lm") + 
  theme(axis.text = element_text(size = 10)) + 
  facet_wrap(~subject)
```

# Model
## linear {.tabset} 

### block  
```{r}
null_m <- lmer(trial_looking_time ~ 1 + (1|subject), 
               data = d)

basic_m <- lmer(trial_looking_time ~ block + (1|subject), 
     data = d)

summary(basic_m)
anova(null_m, basic_m)
```

### stimuli repetition
```{r}
null_m <- lmer(log(trial_looking_time) ~ 1 + (1|subject), 
               data = excluded_sum)

rep_m <- lmer(log(trial_looking_time) ~ num_times_stimulus_seen + (1|subject), 
     data = excluded_sum)

summary(rep_m)
anova(null_m, rep_m)
```

### stimuli repetition and block interaction
```{r}
interaction_m <- lmer(log(trial_looking_time) ~ num_times_stimulus_seen * block + (1|subject), 
     data = excluded_sum)

summary(interaction_m)
anova(null_m, interaction_m)
```



## GAM {.tabset}

only focusing on looking time at the background 
```{r}
library(mgcv)
gam_d <- full_aggregated %>% 
  filter(trial_stimulus_type == "background") %>% 
  mutate(trial_stimulus_complexity = as.factor(trial_stimulus_complexity), 
         block = as.factor(block))
```
### number of repetition
```{r}
gam_m <- gam(trial_looking_time ~ s(num_times_stimulus_seen), 
             data = gam_d, 
             method = "REML")
summary(gam_m)
plot(gam_m, 
     pages = 1, 
     se = TRUE, 
     shade = TRUE)
gam.check(gam_m)
concurvity(gam_m, full = TRUE)
```
### number of repetition (with log?)

```{r}
gam_m <- gam(log(trial_looking_time) ~ s(num_times_stimulus_seen), 
             data = gam_d, 
             method = "REML")
summary(gam_m)
plot(gam_m, 
     pages = 1, 
     se = TRUE, 
     shade = TRUE)
gam.check(gam_m)
concurvity(gam_m, full = TRUE)
```

### number of repetition by stimulus complexity 
```{r fig.width = 15, fig.height = 15}
gam_m <- gam(trial_looking_time ~ s(num_times_stimulus_seen, 
                                    by = trial_stimulus_complexity), 
             data = gam_d, 
             method = "REML")

summary(gam_m)

plot(gam_m, 
     pages = 1, 
     se = TRUE, 
     shade = TRUE)
gam.check(gam_m)
concurvity(gam_m, full = TRUE)
```
### number of repetition by stimulus complexity (log?)
```{r fig.width = 15, fig.height = 15}
gam_m <- gam(log(trial_looking_time) ~ s(num_times_stimulus_seen, 
                                    by = trial_stimulus_complexity), 
             data = gam_d, 
             method = "REML")

summary(gam_m)

plot(gam_m, 
     pages = 1, 
     se = TRUE, 
     shade = TRUE)
gam.check(gam_m)
concurvity(gam_m, full = TRUE)
```
### number of repetition by block 
```{r}
gam_m <- gam(trial_looking_time ~ s(num_times_stimulus_seen, by = block), 
             data = gam_d, 
             method = "REML")

plot(gam_m, all.terms = TRUE, pages = 1,
     shade = TRUE,
     rug = TRUE, 
     se = TRUE)
summary(gam_m)
gam.check(gam_m)
concurvity(gam_m, full = TRUE)
```



### number of repetition by block (log?)
```{r}
gam_m <- gam(log(trial_looking_time) ~ s(num_times_stimulus_seen, by = block), 
             data = gam_d, 
             method = "REML")

plot(gam_m, all.terms = TRUE, pages = 1,
     shade = TRUE,
     rug = TRUE, 
     se = TRUE)
summary(gam_m)

gam.check(gam_m)
concurvity(gam_m, full = TRUE)
```

within subject difference 







