---
title: "Power analysis through simulation for Daffner et al (2000)"
author: "Anjie Cao (anjiecao@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: no
---


```{r include=FALSE}
library(tidyverse)
library(knitr)
library(rstatix)
library(lme4)
library(DT)
library(kableExtra)
library(here)
SIMULATION_NUM = 1000
IMAGE_PATH <- here("original_plot.png")
```

# Credit:
Many code chunks are adapted from the recitation notebook created for [CMU 85732](https://github.com/CoAxLab/DataSciencePsychNeuro_CMU85732/blob/master/RecitationNotebooks/Recitation10_power_analysis_through_simulation.ipynb) 

# Set up the mean and SD as measured from the plot:

The original plot
![original plot](`r IMAGE_PATH`)


```{r}

N <- 24
deviant_simple_mean <- (85.9 / 31.59) * 500
deviant_unusual_mean <- (181.44 / 31.59) * 500
deviant_mixed_mean <- (98.88 / 31.59) * 500

deviant_simple_sem <- (9.92 / 31.59) * 500
deviant_unusual_sem <- (26.1 / 31.59) * 500
deviant_mixed_sem <- (15.6 / 31.59) * 500

deviant_simple_sd <- deviant_simple_sem * sqrt(N)
deviant_unusual_sd <- deviant_unusual_sem * sqrt(N)
deviant_mixed_sd <- deviant_mixed_sem * sqrt(N)

bkgd_simple_mean <- (48.02 / 31.59) * 500
bkgd_unusual_mean <- (53.6 / 31.59) * 500
bkgd_mixed_mean <- (58.53 / 31.59) * 500

bkgd_simple_sem <- (10.52 / 31.59) * 500
bkgd_unusual_sem <- (4.27 / 31.59) * 500
bkgd_mixed_sem <- (16.94 / 31.59) * 500

bkgd_simple_sd <- bkgd_simple_sem * sqrt(N)
bkgd_unusual_sd <- bkgd_unusual_sem * sqrt(N)
bkgd_mixed_sd <- bkgd_mixed_sem * sqrt(N)

```

# Generate artificial data
Assuming normal distribution of the looking time data will lead to some "negative" looking time data.
Also, because the plot only provides data aggregated across participant levels, the generated data is not going to be perfectly fit for lmer analysis where trial-level data is needed. 
```{r}
simulate_data <- function(sample_size, 
                          deviant_simple_mu, deviant_simple_sd,
                          deviant_unusual_mu, deviant_unusual_sd,
                          deviant_mixed_mu, deviant_mixed_sd,
                          bkgd_simple_mu, bkgd_simple_sd,
                          bkgd_unusual_mu, bkgd_unusual_sd,
                          bkgd_mixed_mu, bkgd_mixed_sd){
  
  deviant_simple <- rnorm(sample_size, deviant_simple_mu, deviant_simple_sd)
  deviant_unusual <- rnorm(sample_size,deviant_unusual_mu, deviant_unusual_sd)
  deviant_mixed <- rnorm(sample_size,deviant_mixed_mu, deviant_mixed_sd)
  
  bkgd_simple <- rnorm(sample_size,bkgd_simple_mu, bkgd_simple_sd)
  bkgd_unusual <- rnorm(sample_size,bkgd_unusual_mu, bkgd_unusual_sd)
  bkgd_mixed <- rnorm(sample_size,bkgd_mixed_mu, bkgd_mixed_sd)
  
  raw_data <- data.frame(
    deviant_simple = deviant_simple,
    deviant_unusual = deviant_unusual,
    deviant_mixed = deviant_mixed,
    bkgd_simple = bkgd_simple, 
    bkgd_unusual = bkgd_unusual, 
    bkgd_mixed = bkgd_mixed
  )
  
  raw_data$subject <- seq_len(nrow(raw_data))
 tidy_data <- raw_data %>% 
  pivot_longer(cols = deviant_simple:bkgd_mixed,
               names_to = "trial_info",
               values_to = "trial_looking_time") %>% 
  mutate(
    block = case_when(
      grepl("simple", trial_info) ~ "simple",
      grepl("unusual", trial_info) ~ "unusual", 
      grepl("mixed", trial_info) ~ "mixed"
    ),
    trial_stimulus_type = case_when(
      grepl("deviant", trial_info) ~ "deviant", 
      grepl("bkgd", trial_info) ~ "background"
    )
  ) %>% 
  select(-trial_info)
 
 return(tidy_data)

}




tidy_data <- simulate_data(24,
              deviant_simple_mean, deviant_simple_sd,
              deviant_unusual_mean, deviant_unusual_sd,
              deviant_mixed_mean, deviant_mixed_sd,
              bkgd_simple_mean, bkgd_simple_sd,
              bkgd_unusual_mean, bkgd_unusual_sd,
              bkgd_mixed_mean, bkgd_mixed_sd)

tidy_data %>% datatable()
```





# Analysis functions

## Function to run analysis once
One to run anova, extracting p value. 
The other to fit lmer model, exctracting BIC. 

```{r}
# get the interaction effect p value
run_anova <- function(data){
  
  anova_res <- anova_test(data = data, 
           formula = trial_looking_time ~ trial_stimulus_type * block,
           within = c(trial_stimulus_type,block),
           dv = trial_looking_time)
  # get the p val for the interaction term 
  return(anova_res$p[nrow(anova_res)])
  
}

run_lme <- function(data){
  
  null_res <- lmer(log(trial_looking_time) ~ 1 + (1|subject), 
                   data = data)
  lm_res <- lmer(log(trial_looking_time) ~ trial_stimulus_type * block + (1|subject), 
     data = data)
  
  m_bic <- BIC(null_res, lm_res)$BIC
  stats <- diff(m_bic)
  return(stats)
  
}

```

## Function to run analysis repeatedly
analysis_fun is one of the two functions listed above 
analysis_type is either "anova" or "lmer

```{r message=FALSE, warning=FALSE}
repeat_analysis <- function(n_simulations, alpha, analysis_fun, analysis_type,
                          n_subjects, 
                          deviant_simple_mu, deviant_simple_sd,
                          deviant_unusual_mu, deviant_unusual_sd,
                          deviant_mixed_mu, deviant_mixed_sd,
                          bkgd_simple_mu, bkgd_simple_sd,
                          bkgd_unusual_mu, bkgd_unusual_sd,
                          bkgd_mixed_mu, bkgd_mixed_sd) {
    stat_results <- c() # empty vector to store delta BICs from each simulation
    # loop for repeating the simulation
    for (i in 1:n_simulations) {
        data <- simulate_data(n_subjects, 
                          deviant_simple_mu, deviant_simple_sd,
                          deviant_unusual_mu, deviant_unusual_sd,
                          deviant_mixed_mu, deviant_mixed_sd,
                          bkgd_simple_mu, bkgd_simple_sd,
                          bkgd_unusual_mu, bkgd_unusual_sd,
                          bkgd_mixed_mu, bkgd_mixed_sd)
        stat_result <- analysis_fun(data)
        stat_results <- c(stat_results,stat_result) 
    }
    
    if (analysis_type == "lmer"){
    # calculate how many of the simulations had significant results
    power <- mean(stat_results <= -20) #strong evidence, see Raftery & Kass 1995
    return(list(power = power, stat_results = stat_results))
    } else if(analysis_type == "anova"){
    power <- mean(stat_results <= alpha)
    return(list(power = power, stat_results = stat_results))
    }
}

```

# Simulation results:
The current RPub run `r SIMULATION_NUM` simulations. 

## Anova test
```{r message=FALSE, warning=FALSE}

anova_data <- expand.grid(sample_size = c(10,20,30,40), alpha = c(0.05,0.01,0.001)) 
anova_data$id <- 1:nrow(anova_data) 

anova_results <- anova_data  %>%  
    nest(-id, .key = 'parameters')  %>% 
    mutate(power = map(parameters, ~repeat_analysis(SIMULATION_NUM, .$alpha, run_anova, "anova",
                                                     
                                                     .$sample_size,
                                                     deviant_simple_mean, deviant_simple_sd,
                                                     deviant_unusual_mean, deviant_unusual_sd,
                                                      deviant_mixed_mean, deviant_mixed_sd,
                                                      bkgd_simple_mean, bkgd_simple_sd,
                                                      bkgd_unusual_mean, bkgd_unusual_sd,
                                                      bkgd_mixed_mean, bkgd_mixed_sd)$power)) %>% 
    unnest(parameters, power)





```


```{r}
anova_results %>% 
  ggplot(aes(sample_size, power, color = as.factor(alpha), group = alpha)) +
    geom_point() +
    geom_line() +
    geom_hline(yintercept = 0.8) + 
    geom_hline(yintercept = 0.95) +
    scale_color_discrete('Alpha level') +
    scale_x_continuous('Sample size') +
    theme_classic()
```





## Linear mixed effect model results
```{r message=FALSE, warning=FALSE}
dat <- expand.grid(sample_size = c(10,20,30,40))
dat$id <- 1:nrow(dat)
lmer_results <- dat  %>% 
    nest(-id, .key = 'parameters')  %>% 
    mutate(power = map(parameters, ~ repeat_analysis(SIMULATION_NUM, 0.05, run_lme, "lmer",
                                                     .$sample_size,
                                                     deviant_simple_mean, deviant_simple_sd,
                                                     deviant_unusual_mean, deviant_unusual_sd,
                                                      deviant_mixed_mean, deviant_mixed_sd,
                                                      bkgd_simple_mean, bkgd_simple_sd,
                                                      bkgd_unusual_mean, bkgd_unusual_sd,
                                                      bkgd_mixed_mean, bkgd_mixed_sd)$power))  %>% 
    unnest(parameters, power)

```

```{r}
lmer_results %>% 
  ggplot(aes(sample_size, power)) +
    geom_point() +
    geom_line() +
    geom_hline(yintercept = 0.8) + 
    geom_hline(yintercept = 0.95) +
    scale_color_discrete('Alpha level') +
    scale_x_continuous('Sample size') +
    theme_classic()
```





















