---
title: "Data Preprocessing and Excluding"
author: "anjie"
date: "10/14/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(DT)
library(jsonlite)
library(tidyverse)
library(kableExtra)
RAW_DATA_DIR <- here("raw_data/")
AGGREGATED_DATA_PATH <- here("processed_data/aggregated_data.csv")
```

# Aggregate the Raw Data 
```{r message=FALSE, warning=FALSE}
raw_df <- list.files(RAW_DATA_DIR, full.names = T) %>%
  map_df(read_csv) %>% 
  select(rt, trial_type, trial_index, subject, trial_stimulus, trial_stimulus_type, 
         task_type, task_target_stimulus, task_background_stimulus, task_deviant_stimuli, task_order_number, block_order_number, trial_looking_time, 
         trial_pressed_space_bar,trial_space_bar_rt, responses, question_order) %>% 
  filter(trial_type == "stimulus-presentation" | trial_type == "demog-age" | trial_type == "demog-gender-and-education" | trial_type == "demog-ethnic-US" | trial_type == "demog-disorder-history") %>% 
  mutate(
    stimuli_type = case_when(
    grepl("complex", trial_stimulus) ~ "complex", 
    grepl("simple", trial_stimulus) ~ "simple"
  ))

write_csv(raw_df, AGGREGATED_DATA_PATH)
```

```{r}
raw_df %>% datatable()
```

# Extract the Demog Information 
```{r message=FALSE, warning=FALSE}
demog_df <- raw_df %>% 
  filter(grepl("demog", trial_type)) %>% 
  select(subject, trial_type, responses) %>% 
  toJSON() %>% 
  fromJSON() %>% 
    mutate(
      demog_question = map(responses, ~ fromJSON(.) %>% as.data.frame())) %>% 
    unnest(demog_question)  %>% 
  group_by(subject) %>%
  mutate_at(vars(-group_cols()), function(x) {x[!is.na(x)][1]}) %>%
  distinct() %>% 
  select(-trial_type, responses)
  


demog_df %>% datatable()
```
joining back
```{r message=FALSE, warning=FALSE}
data_with_demog <- left_join(raw_df,demog_df, by = "subject")
data_with_demog %>% datatable()
```


# Exclusion criteria applied 

## filter out based on demog (neurological disorder question)
```{r}
demog_filtered <- data_with_demog %>% 
  filter(
    current_neuro != "Yes" && past_neuro != "Yes" && current_ld != "Yes" && past_ld != "Yes"
  )

demog_filtered
```

## participant-level exclusion: pressing space bar

```{r}
## proportion of pressing space bar in target trial 
prop_press_space_summary <- data_with_demog %>% 
  select(subject, trial_pressed_space_bar, trial_stimulus_type) %>% 
  mutate(trial_pressed_space_bar = if_else(is.na(trial_pressed_space_bar), "no", trial_pressed_space_bar)) %>% 
  filter(trial_stimulus_type == "target") %>% 
  group_by(subject,trial_pressed_space_bar) %>% 
  summarize(
    n = n()
  ) %>% 
  pivot_wider(names_from = trial_pressed_space_bar, values_from = n) %>% 
  mutate(
    no = if_else(is.na(no), 0.0, as.numeric(no)),
    yes = if_else(is.na(yes), 0.0, as.numeric(yes)), 
    sum = no + yes, 
    no_press_prop = no / sum
  ) %>% 
  select(subject, no_press_prop)

## proportion of pressing space bar in non-target trial 
prop_wrong_press_summary <- data_with_demog %>% 
  select(subject, trial_pressed_space_bar, trial_stimulus_type) %>% 
  mutate(trial_pressed_space_bar = if_else(is.na(trial_pressed_space_bar), "no", trial_pressed_space_bar)) %>% 
  filter(trial_stimulus_type != "target") %>% 
  group_by(subject,trial_pressed_space_bar) %>% 
  summarize(
    n = n()
  ) %>% 
  pivot_wider(names_from = trial_pressed_space_bar, values_from = n) %>% 
  mutate(
    no = if_else(is.na(no), 0.0, as.numeric(no)),
    yes = if_else(is.na(yes), 0.0, as.numeric(yes)), 
    sum = no + yes, 
    wrong_press_prop = yes / sum
  ) %>% 
  select(subject, wrong_press_prop)

prop_wrong_press_summary

data_with_demog_spacebar <- left_join(data_with_demog, prop_press_space_summary, by = "subject")
data_with_demog_spacebar <- left_join(data_with_demog_spacebar, prop_wrong_press_summary, by = "subject")

filtered_spacebar <- data_with_demog_spacebar %>% 
  filter(no_press_prop > 0.25 | wrong_press_prop >0.25)
```

## trial-level exclusion:????
```{r}
summary_lt <- data_with_demog %>% 
  filter(trial_type == "stimulus-presentation") %>% 
    summarize(
    mean_lt = mean(as.numeric(trial_looking_time), na.rm = TRUE), 
    sd_lt = sd(trial_looking_time, na.rm = TRUE),
    upper_lt = mean_lt + 3 * sd_lt, 
    lower_lt = mean_lt - 3 * sd_lt
  )

UPPER_LT <- summary_lt %>% select(upper_lt) %>% pull()
LOWER_LT <- summary_lt %>% select(lower_lt) %>% pull()
summary_lt %>% kable()
```

```{r}
trial_lt_to_cut <- data_with_demog %>% 
  filter(trial_type == "stimulus-presentation") %>% 
  filter(trial_looking_time < LOWER_LT | trial_looking_time > UPPER_LT)

# check if the participants are missing 25% trials after applying the trial-based exclusion criteria 
num_total_trials <- data_with_demog %>% 
  filter(trial_type == "stimulus-presentation") %>% 
  group_by(subject) %>% 
  count()

num_to_cut_trials <- trial_lt_to_cut %>% 
  filter(trial_type == "stimulus-presentation") %>% 
  group_by(subject) %>% 
  count()

num_total_trials %>% kable()
num_to_cut_trials %>% kable()
```

transformation needed? 
```{r}
data_with_demog %>% 
  ggplot(aes(x = as.numeric(trial_looking_time))) + 
  geom_histogram() + 
  geom_vline(xintercept = UPPER_LT)
```

# plots 

```{r}
data_with_demog %>% 
  filter(trial_type == "stimulus-presentation") %>% 
  group_by(trial_stimulus_type, task_type) %>% 
  summarize(
    mean = mean(as.numeric(trial_looking_time), na.rm = TRUE),
    sd = sd(as.numeric(trial_looking_time), na.rm = TRUE), 
    n = n(),
    ci_range_95 =  qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)),
    ci_lower = mean - ci_range_95,
    ci_upper = mean + ci_range_95
  ) %>% 
  ggplot(aes(x = trial_stimulus_type, y = mean)) + 
  geom_point() + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~task_type)
```

orirignal plot?
```{r}
task_type_order <- c("all_simple", "all_complex", "mixed_simple_deviant", "mixed_complex_deviant")

data_with_demog %>% 
  filter(trial_type == "stimulus-presentation") %>% 
  filter(trial_stimulus_type != "target") %>% 
  group_by(trial_stimulus_type, task_type) %>% 
  summarize(
    mean = mean(as.numeric(rt), na.rm = TRUE),
    sd = sd(as.numeric(rt), na.rm = TRUE), 
    n = n(),
    ci_range_95 =  qt(1 - (0.05 / 2), n - 1) * (sd/sqrt(n)),
    ci_lower = mean - ci_range_95,
    ci_upper = mean + ci_range_95
  ) %>% 
  ggplot(aes(x = task_type, y = mean)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~trial_stimulus_type) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```





# final clean data 
(it's a lie right now because not excluding anyone but it should look sth similar to this)
```{r}
clean_data <- data_with_demog_spacebar %>% 
  filter(trial_type == "stimulus-presentation") %>% 
  select(subject,  task_type, task_order_number, trial_stimulus_type, trial_stimulus, stimuli_type,
         trial_looking_time, trial_pressed_space_bar, trial_space_bar_rt, age, ethnicity, gender, education, no_press_prop, wrong_press_prop) %>% 
  mutate(
    trial_stimulus_complexity = stimuli_type, 
    trial_stimulus_path = gsub("<img src='", "", trial_stimulus), 
    trial_stimulus_path = gsub("' width ='500' height = '500' style='border:5px solid black'>", "", trial_stimulus_path), 
    demog_age = age, 
    demog_ethnicity = ethnicity, 
    demog_gender = gender, 
    demog_education = education, 
    target_no_press_percent = no_press_prop,
    non_target_press_percent = wrong_press_prop,
    trial_pressed_space_bar = if_else(is.na(trial_pressed_space_bar), "no", trial_pressed_space_bar)
  ) %>% 
  select(
    subject, task_type, task_order_number, 
    trial_stimulus_path, trial_stimulus_type, trial_stimulus_complexity, trial_looking_time, 
    trial_pressed_space_bar, trial_space_bar_rt,  target_no_press_percent, non_target_press_percent, 
    demog_age, demog_ethnicity, demog_gender, demog_education
  )
  
clean_data %>% datatable()
```



```{r}
library(shape)
vignette("shape")
```

